Notes on conformal manifolds

### 2025-07-23 Draft_MLFlatDirections.tex

\section{Supergravity setup}
%\begin{itemize}[label=\textbullet]
%	\item Context: 6d ${\rm AdS}_{3}\times S^{3}$, truncation to 3d, potential, conformal manifold.

Three-dimensional $\mathcal{N}=8$ (half-maximal) gauged supergravity is governed by the Lagrangian
%
\begin{equation}	\label{eq: lagrangian_rephrased} 
	e^{-1}\mathcal{L}=R+\frac1{8}g^{\mu\nu}D_\mu M^{\bM\bN}D_\nu M_{\bM\bN}+e^{-1}\mathcal{L}_{\text{CS}}-V\,,
\end{equation}
%
which comprises the Einstein-Hilbert term $R$, the kinetic term for scalar fields parametrized by the matrix $M_{\bM\bN}$, a Chern-Simons contribution $\mathcal{L}_{\text{CS}}$, and a scalar potential $V$.

@@

Parametrizing the scalar matrix following \cite{Eloy:2021fhc}, we have for the scalar matrix 
\begin{equation}	\label{eq: scalarmatrix}
	M_{\bM\bN}
	=
	\begin{pmatrix}
		m+(\xi^2+\phi)m^{-1}(\xi^2-\phi)+2\xi^2	& 	(\xi^2+\phi)m^{-1}		&	0	&  0  &	-\sqrt2\,[1+(\xi^2+\phi)m^{-1}]\xi	\\
		m^{-1}(\xi^2-\phi)	& 	m^{-1}	&	0  	&  0  &	-\sqrt2\,m^{-1}\xi	\\
		0		&	0		&	e^{2\tilde\varphi}	&	0				&	0		\\
		0		&	0		&	0				&	e^{-2\tilde\varphi}	&	0		\\
		-\sqrt2\,\xi^T[1+m^{-1}(\xi^2-\phi)]	&	-\sqrt2\,\xi^Tm^{-1}	&	0	&	0	&	1+2\,\xi^Tm^{-1}\xi	\\
	\end{pmatrix}\,,
\end{equation}
which encodes 22 scalars of the theory ($22 = 32 - 10$, with 10 scalars gauge fixed using translations in the gauge group), which have been parametrized by 
$m = \nu\nu^{T}\in{\rm GL}(3,\mathbb{R})$ parametrizing the coset ${\rm GL}(3,\mathbb{R})/{\rm SO}(3)$, $\phi$ a $3\times3$ antisymmetric matrix, $\xi$ a $3\times4$ matrix, and $\xi^{2} = \xi \xi^{T}$, a dilaton $\tilde{\varphi}$. With this parametrization, the potential takes the from ref.~\cite{Eloy:2021fhc}:
\begin{equation} \label{eq:scalarpotential}
	\begin{aligned}
		V & = 4\,e^{-4\tilde\varphi}+2\,e^{-2\tilde\varphi}\Big[-\tr\left(m+m^{-1}\right)+\tr\left(\phi m^{-1}\phi\right) -2\,\tr\left(\phi m^{-1}\xi^{2}\right)-2\,\tr\left(\xi^{2}\right)\\
		&\qquad\quad-\tr\left(\xi^{2}m^{-1}\xi^{2}\right)  +\frac{1}{2}\,\det\left(m^{-1}\right)\left(1-\tr\left(\phi^{2}\right)-\tr\left(\xi^{4}\right)+\tr\left(\xi^{2}\right)^{2}\right) \\
		&\qquad\quad +\frac{1}{2}\,{\rm T}\left(m^{-1}(\xi^{2}-\phi),(\xi^{2}+\phi)m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\\
		&\qquad\quad +\frac{1}{4}\,{\rm T}\left(m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\Big], 
	\end{aligned}
\end{equation}
where ${\rm T}\left(A,B,C\right)=\varepsilon_{mnp}\,\varepsilon_{qrs}\,A^{mq}B^{nr}C^{ps}$. We can further restrict ourselves to a set of 13 scalars which is still a consistent truncation, with 
\begin{equation}
	\begin{aligned}
		\xi &= \begin{pmatrix}
					0 & 0 & 0 & x_{1} \\
					0 & 0 & 0 & x_{2} \\
					0 & 0 & 0 & x_{3}
				\end{pmatrix}, \\[5pt]
		\phi &= \begin{pmatrix}
					0 & x_{4} & x_{5} \\
					-x_{4} & 0 & x_{6} \\
					-x_{5} & -x_{6} & 0
				\end{pmatrix}, \\[5pt]
		\tilde{\varphi} &= \tilde{x}_{13}, \\[5pt]
		\nu &= e^{(6\,\tilde{x}_{7}+3\,\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/6}
				\begin{pmatrix}
					1 & \frac{x_{10}}{\sqrt{2}} & \frac{x_{11}}{\sqrt{2}} + \frac{x_{10}x_{12}}{4} \\
					0 & e^{-\tilde{x}_{8}} & \frac{e^{-\tilde{x}_{8}}\,x_{12}}{\sqrt{2}} \\
					0 & 0 & e^{-(\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/2}
				\end{pmatrix}.
	\end{aligned}
\end{equation}

We note here that all dilaton fields are denoted with a tilde. This convention is adopted in anticipation of a later redefinition of the form \( x_i = e^{\tilde{x}_i} \) for these fields. By reserving the notation without tildes for the exponentiated variables, we maintain a clear and consistent distinction for use in subsequent expressions.

A first numerical exploration suggests that attention can be focused on the five scalars $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x_{10}$. This subset, selected at random from the original set of thirteen scalar fields, serves to render the problem more tractable at the outset. It is important to emphasize, however, that this subset does not constitute a consistent truncation of the full theory. Consequently, the truncated theory involving only five scalar fields may not constitute a formally consistent truncation of the full 13-scalar theory, as higher-order interactions could, in principle, source the remaining scalar modes. As a result, the interpretation of the five-scalar theory as a consistent subsector of the original theory may no longer hold. Nonetheless, the five-scalar model remains a self-consistent theory in its own right and can be used independently. This distinction does not affect our numerical approach.

Nevertheless, the solutions we obtain remain valid solutions of the complete theory. This is because we identify the critical points of the scalar potential by first computing $\nabla V = 0$ with all scalar fields included, and only then setting the remaining fields to zero. By performing the differentiation before the truncation, we ensure that the resulting configurations satisfy the full equations of motion and are therefore legitimate solutions of the complete theory.

To compute the gradient, we exploit the fact that we possess an analytic expression for the scalar potential \eqref{eq:scalarpotential}, which we implement in Mathematica for numerical analysis. Note however that it could have been fully done using a numerical approach, automatic differenciations methods (that we will use later to do the gradient descent for the loss function).
%Under this truncation, the scalar potential~\eqref{eq:scalarpotential} reduces to:
%\begin{equation} \label{eq:scalarpotential124810}
%	\begin{aligned}
%		V &= \frac{1}{8}\, e^{-2\,\tilde{x}_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,\tilde{x}_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
%		  & \quad\qquad\qquad - 4\,e^{3\,\tilde{x}_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
%      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
%        & \quad\qquad\qquad - 8\,e^{\tilde{x}_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
%      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
%        & \quad\qquad\qquad + 4\,e^{2\,\tilde{x}_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
%        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
%        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,\tilde{x}_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
%	\end{aligned}
%\end{equation}

@@

An additional advantage of this approach is its broader applicability. In scenarios where an explicit analytic expression for the potential and its gradient is inaccessible—yet numerical evaluation remains feasible—our method retains its validity. In particular, one could replace the analytic gradient by automatic differenciation methods if one can only evaluate the potential. Hence, the numerical strategy adopted here may prove effective even in cases where an exact analytic expression for the potential is out of reach. \bd{other motivation? Example of this ? }

\section{Numerical analysis}
	\subsection{Gradient descent and local analysis}
		To identify the flat directions in the potential, we will use numerical tools. The procedure is as follows: first, we sample the underlying manifold, and then we use numerical techniques to extract analytical information from the resulting cloud of points. The first step is thus to sample the manifold.
			\subsubsection{Sampling the manifold}
		%\paragraph{Gradient descent: implementation, sampling, time needed, loss}

		To do so, we perform a basic gradient descent. We begin by randomly and uniformly initializing points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range $[-1,1]$. This will be important later, when performing the symbolic regression \footnote{The rationale is that symbolic regression (SR) evaluates the data points ${\vec{x}^i}$ using a candidate polynomial function $p$. The loss function—used later to assess the quality of a given polynomial—is defined as $\sum_i p(\vec{x}^i)$. When all data points lie within the interval $[-1, 1]$, the SR algorithm tends to favor high-degree polynomials, as these often yield smaller values over this range, thereby artificially lowering the loss. Introducing data points with absolute values greater than 1 mitigates this bias by discouraging the selection of overly complex polynomials that exploit this effect.}. For this example, we choose to generate $10^5$ points. This number is motivated by the dimensionality of our problem. If we aim to populate all possible directions and want approximately $\mathcal{O}(10)$ points per direction, then we require $\mathcal{O}(10^5)$ points. Given that the true intrinsic dimensionality of the manifold is less than or equal to 5, this value serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

		We then perform gradient descent on the points using TensorFlow’s automatic differentiation \cite{tensorflow2015-whitepaper}. The loss function is defined as:
		\begin{equation} \label{eq:lossgraddes}
			\mathcal{L} = \sum_{i = 1}^{n_{points}} ||\left.\nabla V(\vec{X}^{\,i})\,\right|_{\vec{y}=0}||^2
		\end{equation}
		where $\vec{X}^{\,i} = (x^i_1,\dots,x^i_6,\tilde{x}^i_7,\tilde{x}^i_8,\tilde{x}^i_9,x^i_{10},\dots,x^i_{12},\tilde{x}^i_{13})$ and $\vec{y}^{\,i} = (x^i_3,x^i_5,x^i_6,\tilde{x}^i_7,\tilde{x}^i_9,x^i_{11},x^i_{12},\tilde{x}^i_{11})$ denotes the data points. This way we only keep $(x^i_1,x^i_2,x^i_4,\tilde{x}^i_8,x^i_{10})$ alive after we have taken the derivative. As already mentionned above, we use the analytic formula for $\nabla V$, but we could have perform the gradient descent by fully using automatic differentiation.

		Through trial and error, we observed that periodically reinitializing the optimizer significantly improved the convergence rate. This technique bears similarity to the concept of warm restarts introduced in \cite{loshchilov2017sgdrstochasticgradientdescent}. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialized, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.  The evolution of the loss function, along with the learning rate schedule, is shown in Fig.~\eqref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is reinitialized. We also observe that, following the last few reinitializations, the loss exhibits a small bump immediately after the restart. We interpret this behavior as a consequence of the increased learning rate: some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. This strategy introduces the risk of desampling certain regions in favor of others, introducing the risk to hide some flat directions in the potential. However, we believe that with enough points, it is very unlikely for a flat direction to be completely hidden from us and desampled.

		\begin{figure}
			\centering
			\subfigure{ \label{fig:log_grad_V}
				\begin{tikzpicture}
			  		\draw (0,0) node (fig1) {\includegraphics[scale=0.7]{Figures/log_grad_V_hist_21_05_2025.pdf}};
			  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
			  		\draw (9,0) node (fig2) {\includegraphics[scale=0.7]{Figures/log_Vp4_hist_21_05_2025.pdf}};
			  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
			  	\end{tikzpicture}
			}
			\subfigure{\label{fig:log_Vp4}}
			\vspace{-1cm}
			\caption{(a) Histogram of the the norm of the gradient for each point ($x$ axis is in $\log$ scale). (b) Histogram of the distance of the value of the potential to $V(0)$ for each point ($x$ axis is in $\log$ scale).}
			\label{fig:Analyze_V_points}
		\end{figure}

		The Adam optimizer was employed throughout the gradient descent procedure. It was reinitialized at iterations {250, 500, 750}, with the learning rate fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialized once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialization was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs.

		Upon completion, it was observed that all data points converged to values close to $-4$. Specifically, nearly all data points attained values within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were on the order of $10^{-3}$. Similarly, 99,739 out of 100,000 data points exhibited gradient norms smaller than $10^{-3}$. These results are summarized in Figure~\ref{fig:Analyze_V_points}.

		\begin{figure}[h!]
			\centering
			\includegraphics[scale = 0.75]{Figures/Loss_grad_des_21_05_2025.pdf}
			\caption{Evolution of the loss function~\eqref{eq:lossgraddes} during gradient descent and corresponding learning rate schedule, noted $\alpha$ on the left y-axis, both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
			\label{fig:loss_grad_des}
		\end{figure}

>>>>

\section{Supergravity setup} \label{sec:sugra}
Three-dimensional $\mathcal{N}=8$ (half-maximal) gauged supergravity is governed by the Lagrangian
%
\begin{equation}	\label{eq: lagrangian_rephrased} 
	e^{-1}\mathcal{L}=R+\frac1{8}g^{\mu\nu}D_\mu M^{\bM\bN}D_\nu M_{\bM\bN}+e^{-1}\mathcal{L}_{\text{CS}}-V\,,
\end{equation}
%
which comprises the Einstein-Hilbert term $R$, the kinetic term for scalar fields parametrized by the matrix $M_{\bM\bN}$, a Chern-Simons contribution $\mathcal{L}_{\text{CS}}$, and a scalar potential $V$.

@@

Parametrizing the scalar matrix following \cite{Eloy:2021fhc}, we have for the scalar matrix 
\begin{equation}	\label{eq: scalarmatrix}
	M_{\bM\bN}
	=
	\begin{pmatrix}
		m+(\xi^2+\phi)m^{-1}(\xi^2-\phi)+2\xi^2	& 	(\xi^2+\phi)m^{-1}		&	0	&  0  &	-\sqrt2\,[1+(\xi^2+\phi)m^{-1}]\xi	\\
		m^{-1}(\xi^2-\phi)	& 	m^{-1}	&	0  	&  0  &	-\sqrt2\,m^{-1}\xi	\\
		0		&	0		&	e^{2\tilde\varphi}	&	0				&	0		\\
		0		&	0		&	0				&	e^{-2\tilde\varphi}	&	0		\\
		-\sqrt2\,\xi^T[1+m^{-1}(\xi^2-\phi)]	&	-\sqrt2\,\xi^Tm^{-1}	&	0	&	0	&	1+2\,\xi^Tm^{-1}\xi	\\
	\end{pmatrix}\,,
\end{equation}
which encodes 22 scalars of the theory ($22 = 32 - 10$, with 10 scalars gauge fixed using translations in the gauge group), which have been parametrized by 
$m = \nu\nu^{T}\in{\rm GL}(3,\mathbb{R})$ parametrizing the coset ${\rm GL}(3,\mathbb{R})/{\rm SO}(3)$, $\phi$ a $3\times3$ antisymmetric matrix, $\xi$ a $3\times4$ matrix, and $\xi^{2} = \xi \xi^{T}$, a dilaton $\tilde{\varphi}$. With this parametrization, the potential takes the from ref.~\cite{Eloy:2021fhc}:
\begin{equation} \label{eq:scalarpotential}
	\begin{aligned}
		V & = 4\,e^{-4\tilde\varphi}+2\,e^{-2\tilde\varphi}\Big[-\tr\left(m+m^{-1}\right)+\tr\left(\phi m^{-1}\phi\right) -2\,\tr\left(\phi m^{-1}\xi^{2}\right)-2\,\tr\left(\xi^{2}\right)\\
		&\qquad\quad-\tr\left(\xi^{2}m^{-1}\xi^{2}\right)  +\frac{1}{2}\,\det\left(m^{-1}\right)\left(1-\tr\left(\phi^{2}\right)-\tr\left(\xi^{4}\right)+\tr\left(\xi^{2}\right)^{2}\right) \\
		&\qquad\quad +\frac{1}{2}\,{\rm T}\left(m^{-1}(\xi^{2}-\phi),(\xi^{2}+\phi)m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\\
		&\qquad\quad +\frac{1}{4}\,{\rm T}\left(m^{-1},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2},m+(\xi^{2}+\phi)m^{-1}(\xi^{2}-\phi)+2\,\xi^{2}\right)\Big], 
	\end{aligned}
\end{equation}
where ${\rm T}\left(A,B,C\right)=\varepsilon_{mnp}\,\varepsilon_{qrs}\,A^{mq}B^{nr}C^{ps}$. We can further restrict ourselves to a set of 13 scalars which is still a consistent truncation, with 
\begin{equation}
	\begin{aligned}
		\xi &= \begin{pmatrix}
					0 & 0 & 0 & x_{1} \\
					0 & 0 & 0 & x_{2} \\
					0 & 0 & 0 & x_{3}
				\end{pmatrix}, \\[5pt]
		\phi &= \begin{pmatrix}
					0 & x_{4} & x_{5} \\
					-x_{4} & 0 & x_{6} \\
					-x_{5} & -x_{6} & 0
				\end{pmatrix}, \\[5pt]
		\tilde{\varphi} &= \tilde{x}_{13}, \\[5pt]
		\nu &= e^{(6\,\tilde{x}_{7}+3\,\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/6}
				\begin{pmatrix}
					1 & \frac{x_{10}}{\sqrt{2}} & \frac{x_{11}}{\sqrt{2}} + \frac{x_{10}x_{12}}{4} \\
					0 & e^{-\tilde{x}_{8}} & \frac{e^{-\tilde{x}_{8}}\,x_{12}}{\sqrt{2}} \\
					0 & 0 & e^{-(\tilde{x}_{8}+\sqrt{3}\,\tilde{x}_{9})/2}
				\end{pmatrix}.
	\end{aligned}
\end{equation}
For later convenience, we define a thirteen-dimensional vector
\begin{equation} \label{eq:defvecX}
	% \vec{X} = (x_1,\dots,x_6,\tilde{x}_7,\tilde{x}_8,\tilde{x}_9,x_{10},\dots,x_{12},\tilde{x}_{13})
	\vec{X} = (x_1, x_{2}, x_{3}, x_{4}, x_{5}, x_6, \tilde{x}_7, \tilde{x}_8, \tilde{x}_9,x_{10}, x_{11}, x_{12}, \tilde{x}_{13})
\end{equation}
encompassing all parameters. Note here that all dilaton fields are denoted with a tilde. This convention is adopted in anticipation of a later redefinition of the form \( x_i = e^{\tilde{x}_i} \) for these fields. %By reserving the notation without tildes for the exponentiated variables, we maintain a clear and consistent distinction for use in subsequent expressions.

A first numerical exploration suggests that attention can be focused on the five scalars $x_{1}, x_{2}, x_{4}, \tilde{x}_{8}$, and $x_{10}$. This subset, selected at random from the original set of thirteen scalar fields, serves to render the problem more tractable at the outset. It is important to emphasize, however, that this subset does not constitute a consistent truncation of the full theory. Consequently, the truncated theory involving only five scalar fields may not constitute a formally consistent truncation of the full thirteen-scalar theory, as higher-order interactions could, in principle, source the remaining scalar modes. As a result, the interpretation of the five-scalar theory as a consistent subsector of the original theory may no longer hold. Nonetheless, the five-scalar model remains a self-consistent theory in its own right and can be used independently. This distinction does not affect our numerical approach: the solutions we obtain remain valid solutions of the complete theory. This is because we identify the critical points of the scalar potential by first computing $\nabla V = 0$ with all scalar fields included, and only then setting the remaining fields
\begin{equation} \label{eq:defvecy}
	\vec{y} = (x_3,x_5,x_6,\tilde{x}_7,\tilde{x}_9,x_{11},x_{12},\tilde{x}_{11})
\end{equation}
to zero, as detailed in sec.~\ref{sec:graddes}. By performing the differentiation before the truncation, we ensure that the resulting configurations satisfy the full equations of motion and are therefore legitimate solutions of the complete theory.

To compute the gradient, we exploit the fact that we possess an analytic expression for the scalar potential~\eqref{eq:scalarpotential}, which we implement in Mathematica for numerical analysis. Note however that it could have been fully done using a numerical approach, automatic differentiations methods (that we will use later to do the gradient descent for the loss function).
%Under this truncation, the scalar potential~\eqref{eq:scalarpotential} reduces to:
%\begin{equation} \label{eq:scalarpotential124810}
%	\begin{aligned}
%		V &= \frac{1}{8}\, e^{-2\,\tilde{x}_{8}} \bigg[4 + 4\,x_{1}^4 + e^{4\,\tilde{x}_{8}} \big(2 + x_{10}^2\big)^2 \big(1 + x_{2}^4\big)\\
%		  & \quad\qquad\qquad - 4\,e^{3\,\tilde{x}_{8}}\,\big(2 + x_{10}^2\big) \Big(2 - x_{1}^2 + \sqrt{2}\,x_{1} x_{10} x_{2}^3 + x_{2}^4 + 
%      			2\,x_{1} x_{2} x_{4} + x_{4}^2 - x_{2}^2 \big(1 - x_{1}^2 + x_{4}^2\big)\Big)\\
%        & \quad\qquad\qquad - 8\,e^{\tilde{x}_{8}}\,\Big(2 + x_{1}^4 + \sqrt{2}\,x_{1}^3 x_{10} x_{2} - x_{2}^2 - 2\,x_{1} x_{2} x_{4} + 
%      			x_{4}^2 - x_{1}^2 \big(1 - x_{2}^2 + x_{4}^2\big)\Big)\\
%        & \quad\qquad\qquad + 4\,e^{2\,\tilde{x}_{8}}\,\Big(2 - 4\,x_{1}^2 + x_{1}^4 - 4\,x_{2}^2 + 4\,x_{1}^2 x_{2}^2 + x_{2}^4
%        			+ x_{10}^2\big(1 + 3\,x_{1}^2 x_{2}^2\big) + 4\,x_{4}^2 + x_{4}^4\Big)\\
%        & \quad\qquad\qquad + 8\sqrt{2}\,x_{10}\,e^{2\,\tilde{x}_{8}}\,\Big(x_{1}^3 x_{2} + x_{1}^2 x_{4} - x_{2}^2 x_{4} + x_{1} \big(x_{2}^3 - x_{2} x_{4}^2\big)\Big)\bigg].
%	\end{aligned}
%\end{equation}

@@

An additional advantage of this approach is its broader applicability. In scenarios where an explicit analytic expression for the potential and its gradient is inaccessible—yet numerical evaluation remains feasible—our method retains its validity. In particular, one could replace the analytic gradient by automatic differentiation methods if one can only evaluate the potential. Hence, the numerical strategy adopted here may prove effective even in cases where an exact analytic expression for the potential is out of reach. \bd{other motivation? Example of this ? }

\section{Numerical analysis}
To identify the flat directions in the potential, we will use numerical tools. The procedure is as follows: first, we sample the underlying manifold, and then we use numerical techniques to extract analytical information from the resulting cloud of points. The sampling of the manifold is done through a gradient descent. We then analyse its dimension and structure using a local Principal Component Analysis~(PCA) and clustering algorithms, and finally extract analytical constraints defining the manifold thanks to Symbolic Regression methods. We detail these different procedures in the following.

	\subsection{Sampling the manifold: gradient descent} \label{sec:graddes}
	To perform the gradient descent, we initialise randomly and uniformly points within a hypercube of range $[-2,2]$. This choice is important to ensure that points are not restricted to the inner range~$[-1,1]$, which will be crucial for the Symbolic Regression.\footnote{The rationale is that symbolic regression (SR) evaluates the data points $\vec{x}^{(i)}$ using a candidate polynomial function $p$. The loss function—used later to assess the quality of a given polynomial—is defined as $\sum_i p(\vec{x}^{(i)}$. When all data points lie within the interval $[-1, 1]$, the SR algorithm tends to favor high-degree polynomials, as these often yield smaller values over this range, thereby artificially lowering the loss. Introducing data points with absolute values greater than 1 mitigates this bias by discouraging the selection of overly complex polynomials that exploit this effect.} \ce{Put the footnote later in the SR section?} As the example we're focussing on has a five-dimensional parameter space, we choose to generate $n_{\rm points}=10^5$ points: we aim at populating all five directions and want approximately $\mathcal{O}(10)$ points per direction. As the solutions manifold lie within the five dimensional space, its intrinsic dimension is less than or equal to 5. The value $10^5$ then serves as a conservative upper bound for the number of points needed to adequately sample the manifold.

	We then perform a gradient descent on the points using TensorFlow’s automatic differentiation \cite{tensorflow2015-whitepaper}. The loss function is defined as
	\begin{equation} \label{eq:lossgraddes}
		\mathcal{L} = \sum_{i = 1}^{n_{\rm points}} \left|\left|\left.\nabla V(\vec{X}^{(i)})\,\right|_{\vec{y}^{\,(i)}=0}\right|\right|^2,
	\end{equation}
	% where $\vec{X}^{(i)} = (x^{(i)}_1,\dots,x^{(i)}_6,\tilde{x}^{(i)}_7,\tilde{x}^{(i)}_8,\tilde{x}^{(i)}_9,x^{(i)}_{10},\dots,x^{(i)}_{12},\tilde{x}^{(i)}_{13})$ and $\vec{y}^{\,(i)} = (x^{(i)}_3,x^{(i)}_5,x^{(i)}_6,\tilde{x}^{(i)}_7,\tilde{x}^{(i)}_9,x^{(i)}_{11},x^{(i)}_{12},\tilde{x}^{(i)}_{11})$
	see eq.~\eqref{eq:defvecX} and~\eqref{eq:defvecy} for the definitions of the vectors $\vec{X}$ and $\vec{y}$. The exponents $(i)$ denote the data points, with $i\in\llbracket1,n_{\rm points}\rrbracket$. This way we only keep $(x^{(i)}_1,x^{(i)}_2,x^{(i)}_4,\tilde{x}^{(i)}_8,x^{(i)}_{10})$ alive after we have taken the derivative. As already mentioned above, we use the analytic formula for $\nabla V$, but we could have performed the gradient descent by fully using automatic differentiation.

	\begin{figure}
		\centering
		\includegraphics[scale = 0.75]{Figures/Loss_grad_des_21_05_2025.pdf}
		\caption{Evolution of the loss function~\eqref{eq:lossgraddes} during gradient descent (right $y$-axis, in blue) and corresponding learning rate schedule $\alpha$ (left $y$-axis in orange), both plotted on a logarithmic scale. Dashed lines indicate epochs at which the optimizer was reinitialized.}
		\label{fig:loss_grad_des}
	\end{figure}

	The Adam optimizer was employed throughout the gradient descent procedure~\cite{kingma2017adammethodstochasticoptimization}. Through trial and error, we observed that periodically reinitialising the optimizer significantly improved the convergence rate. It was reinitialised at iterations {250, 500, 750}, with the learning rate $\alpha$ fixed at $10^{-2}$. At iteration 1000, the optimizer was reinitialised once more, this time with a reduced learning rate of $10^{-3}$. A final reinitialisation was performed at iteration 1500, setting the learning rate to $10^{-4}$, and the optimization was continued for an additional 500 epochs. The evolution of the loss function~\eqref{eq:lossgraddes}, along with the learning rate schedule, is shown in fig.~\ref{fig:loss_grad_des}. As can be seen in this figure, the convergence rate improves significantly each time the optimizer is reinitialised. We also observe that, following the last few reinitialisations, the loss exhibits a small bump immediately after the restart. We interpret this behaviour as follows: the learning rate gets internally adjusted by the Adam optimizer during the descent, and may well be smaller than the instructions when the reset occurs. The learning rate then gets suddenly increased, and some points that previously had low loss values may momentarily worsen before benefiting from faster convergence. This effect is likely due to points initially located in regions with weak attractive basins being pushed toward areas where the potential gradient is steeper, thus accelerating their convergence. This strategy introduces the risk of desampling certain regions in favour of others, introducing the risk to hide some flat directions in the potential. However, we believe that with enough points, it is very unlikely for a flat direction to be completely hidden from us and desampled. This method bears similarity to the concept of warm restarts introduced in ref.~\cite{loshchilov2017sgdrstochasticgradientdescent}, where the authors reset the learning rate to some value at each reset, without reinitialising the whole optimizer. However, we found that simply scheduling the learning rate without resetting the optimizer yielded slower convergence. We interpret this as follows: when the optimizer is reinitialised, it effectively "forgets" its past gradient history. As a result, the actual learning rate used corresponds more closely to the specified value, rather than being internally adjusted based on accumulated past gradients. This effect seems to contribute to faster convergence in our case.

	\begin{figure}
		\centering
		\subfigure{ \label{fig:log_grad_V}
			\begin{tikzpicture}
		  		\draw (0,0) node (fig1) {\includegraphics[scale=0.7]{Figures/log_grad_V_hist_21_05_2025.pdf}};
		  		\draw ($(fig1.north west) + (0,0)$) node {\small (a)};
		  		\draw (9,0) node (fig2) {\includegraphics[scale=0.7]{Figures/log_Vp4_hist_21_05_2025.pdf}};
		  		\draw ($(fig2.north west) + (0,0)$) node {\small (b)};
		  	\end{tikzpicture}
		}
		\subfigure{\label{fig:log_Vp4}}
		\vspace{-1cm}
		\caption{(a) Histogram of the the norm of the gradient for each point ($x$-axis is in $\log$ scale). (b)~Histogram of the distance of the value of the potential to $V(0)$ for each point ($x$-axis is in $\log$ scale). Note that four points have deviations of the order of $10^{-3}$ and are not visible on the figure.}
		\label{fig:Analyze_V_points}
	\end{figure}

	Upon completion, the loss function converges around $0.002$. The distribution of the values of the gradient at each of the data points is plotted in fig.~\ref{fig:log_grad_V}. More than $93\%$ of the points converged to values of $\vert\nabla V\vert$ lower than $10^{-4}$, and more than $99\%$ of them exhibited gradient norms smaller than $10^{-3}$, yielding a satisfying sampling of the flat directions. As can be observed in fig.~\ref{fig:log_Vp4}, all data points converged to values close to $V(0)=-4$. Specifically, nearly all points attained $V(0)$ within an absolute error of at least $10^{-4}$, with the exception of four points whose deviations were of the order of $10^{-3}$ (not visible on the graph). As the origin $\vec{X}=0$ of the parameter space corresponds to the three-dimensional truncation of the round ${\rm AdS}_{3}\times S^{3}$ solution of half-maximal supergravity in six dimensions, this confirms that the data points lie within flat directions of the potential.

	\begin{figure}
		\centering
		\makebox[\textwidth][c]{\includegraphics[scale = 0.7]{Figures/Triangular_plot_1_2_4_8_10_21_05_2025.pdf}}
		\caption{Triangular plot showing all $2d$ projections and $1d$ histograms of the data after the gradient descent.}
		\label{triangular_plot_1_2_4_8_10}
	\end{figure}

	As a first visualisation, we present a triangular plot of the data in fig.~\ref{triangular_plot_1_2_4_8_10}. \ce{Sure of the use of ``triangular''?} This figure shows all possible 2d projections of the data, along with the 1d histograms of each coordinate after gradient descent. Note that all directions appear to be well populated. There also seems to be non-trivial correlations in the data, see for example the $x_1/x_2$, $x_4/\tilde{x}_8$ or $x_4/x_{10}$ graphs. They could be genuine correlations, or may result from larger basins of attraction. Thanks to the analysis of sec.~\ref{sec:results}, we will conclude that the structures in the $x_1/x_2$ plot are artefacts of the gradient descent, whereas the ones seen in the $x_4/\tilde{x}_8$ or $x_4/x_{10}$ plots reflect genuine features of the manifold.
